Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-10 fib2.tex
hopefully approaching a fixed point

We model the system on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
This treatment applies equally well to lattice models, where anyons 
can only be located at a discrete set of points, 
or continuum models, where a discretization of the manifold must 
be chosen in order to perform charge measurements. 
In the former case, each tile of $\Lambda$ represents a 
lattice site, while in the latter case each tile represents 
a measurement region. For more details of an analogous model, see~\cite{Brell2013}.

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this 
system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\vac}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_i$ the projector to charge $\vac$ at site (or tile) $i$, i.e.~the ground 
space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above. We identify this ground space 
as the codespace of our model. Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$. It was 
seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

At some point, with strong enough noise, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
However, for finite lattice size, this regime is somewhere
beyond the decoder threshold, which is the 
point at which the decoder creates nontrivial (extensive) loops.
\stf{CGB and I are still not confident in the previous sentence or the next one}
We can view the system state as a tensor network, on which
the decoder is performing an (efficiently simulable) contraction.
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
\dude{more polish needed...}

@@

In order to keep track of the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra piece of curve to the lattice supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$ and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
\dude{Heisenberg picture?}
The observable corresponding to the tile is then
represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
%\cggb{This paragraph and the next still needs another look.}

>>>>

%We model the system on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
%This treatment applies equally well to lattice models, where anyons 
%can only be located at a discrete set of points, 
%or continuum models, where a discretization of the manifold must 
%be chosen in order to perform charge measurements. 
%In the former case, each tile of $\Lambda$ represents a 
%lattice site, while in the latter case each tile represents 
%a measurement region. For more details of an analogous model, see~\cite{Brell2013}.

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ observables arranged in
a square lattice tiling $\Lambda$.
These \emph{tiles} will be the oservables accesible to the error-correction procedure (\emph{decoder}).
We imagine an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\vac}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_i$ the projector to charge $\vac$ at tile $i$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
\dude{ouch this no longer works..}
Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above.
We identify this ground space as the codespace of our model. 
Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

At some point, with strong enough noise, the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
However, for finite lattice size, we use heuristics
to minimize non-local braid moves, which helps to
maintain the state as a separable product. \dude{???}
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$
%However, for finite lattice size, this regime is somewhere
%beyond the decoder threshold, which is the 
%point at which the decoder creates nontrivial (extensive) loops.
%\stf{CGB and I are still not confident in the previous sentence or the next one}
%We can view the system state as a tensor network, on which
%the decoder is performing an (efficiently simulable) contraction.
%This enables simulation of error correction in
%the area around decoder threshold for lattice sizes up to $L=128.$
%\dude{more polish needed...}

@@

In order to keep track of the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra piece of curve to the lattice supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$ and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
%\dude{Heisenberg picture?}
%The observable corresponding to the tile is then
%represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
%\cggb{This paragraph and the next still needs another look.}
